### 2014-08-27 project-mutual-modeling.tex
Raw import of studies 2->5 from Word

\documentclass[twocolumn]{article}

\usepackage[english]{babel}

\usepackage{graphicx}
\graphicspath{{figs/}}

\usepackage{tikz}
\usetikzlibrary{shapes}

\usepackage{subcaption}

\usepackage[draft]{fixme}

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\al}{{\textit{et al.\ }}}

\newcommand{\M}[3]{{\mathcal{M}(#1, #2, #3)}}
\newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
\newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
\newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}

\newcommand{\groundingcriterion}{{$\mathcal{M}^{\circ}_{min}$}}
\newcommand{\inigrounding}{{$\mathcal{M}^{\circ}_{t_0}$}}

\newcommand{\concept}[1]{{\small \texttt{#1}}}
\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}

\title{Is Partner Modeling Mutual?}

\begin{document}
\maketitle

\begin{abstract}
    TDB
\end{abstract}

\section{Introduction}

From its very beginning, CSCL research has been following Roschelle \& Teasley's
(1986) suggestion that collaborative learning has something to do with the
process of constructing and maintaining a shared understanding of the task at
hand. Building a shared/mutual understanding refers to the upper class of
collaborative learning situations, those in which students should build upon
each other's understanding to refine their own understanding. What is expected
to produce learning is not the mere fact that two students build the same
understanding but the cognitive effort they have to engage to build this shared
understanding (Schwartz, 1995). This effort can be observed by the frequency of
rich interactions, i.e. interactions whose occurrence has been related to
learning: (self-) explanations in cognitive science (Chi ; Webb), conflict
resolution in socio-cognitive theories (Doise), mutual regulation (Blaye,) in a
Vygostkian perspective, etc. The construction of a shared understanding has been
investigated for several years in psycholinguistics, under the  notion of
“grounding” (Clark \&Wilkes-Gibbs, 1986). However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned. The monitoring
and repair of mis-understanding explains for instance referential failures in
short dialogue episodes but does hardly predict conceptual change over longer
sessions (Dillenbourg \& Traum, 2006). The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:
"collaborative learning is associated with the increased cognitive-interactional
effort involved in the transition from learning to understand each other to
learning to understand the meanings of the semiotic tools that constitute the
mediators of interpersonal interaction" (Baker, Hansen, Joiner \& Traum, 1999).
Several scholars suggest that CSCL research should go deeper in the
understanding of how partners engage into shared meaning making (Stahl, 2006) or
'intersubjective' meaning making (Suthers (2006).

Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
change viewpoint, it is criticized at the same time as being too cognitivist by
some psycholinguists, i.e. as overestimating the amount shared knowledge and
mutual representations actually necessary to conduct a dialogue. The fundamental
issue, as old as philosophy, is the degree of coupling between the different
levels of dialogue, mostly between the lexical / syntactical level and the
deeper semantic levels. Pickering \& Garrod (2004) argue that the mutual
understanding starts mostly with a 'local alignment' at the level of the
linguistic representations, due to priming mechanisms, and that this superficial
alignment may – in some cases- lead to a 'global alignment' of the semantic
level.  For these authors, the convergence in dialogue, and even the repair of
some misunderstandings, is explained by this mimetic behavior more than by a
monitoring of each other's knowledge: "…interlocutors do not need to monitor and
develop full common ground as a regular, constant part of routine conversation,
as it would be unnecessary and far too costly. Establishment of full common
ground is, we argue, a specialized and non-automatic process that is used
primarily in times of difficulty (when radical misalignment becomes apparent)."
(p. 179). This view is actually not incompatible with Clark's grounding
criterion (REF): the degree of shared understanding that peers need to reach
depends upon the task they perform. For instance, a dialogue between two
surgeons might rely on superficial alignment if they talk about their friends
but has to guarantee accurate common grounds when talking about which
intervention will be conducted in which way on which patient.  Some CSCL
scholars stressed the 'illusion of shared understanding' (JACCO?), i.e. the
potential decoupling between linguistic alignment and actual shared meanings.

This interesting cognitive science debate mostly occurred outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes. Is linguistic alignment sufficient to trigger conceptual change? Does
negotiation of meaning only occurs when partners monitor and diagnose each
other's knowledge. If the ratio between shallow alignment and deep grounding
depends upon the task, and if deep grounding is a condition for learning, then
the pedagogical challenge is to design tasks that require deep grounding. Most
empirical studies on grounding and alignment are conducted with tasks, which
despite being qualified as 'ecologically valid' by their authors, are mere
referencing tasks such as asking the way to the train station or helping the
peer to choose a picture among many. In this contribution, we explore several
richer tasks such as arguing on a sensitive issue or building a concept map.

Deep grounding or shared meaning making requires some cognitive load. For Clark
\& Wilkes-Gibbs (1986), what is important is not the individual effort made by
the receiver of a communicative act, but the overall least collaborative effort.
The cost of producing a perfect utterance may be higher than the cost of
repairing the problems that may arise through misunderstandings. For instance,
subjects are less careful about adapting their utterances to their partner when
they know they can provide feedback on his/her understanding (Schober, 1993). We
introduced the notion of ‘optimal collaborative effort’ (Dillenbourg et al,
1996) to stress that misunderstanding should not be viewed as something to be
avoided (if this was possible), but as an opportunity to engage into
verbalization, explanation, negotiation, and so forth. This issue is related to
the global argument regarding cognitive load in learning activities, especially
in discovery learning environments: there is no learning without some cognitive
load, but overload may hinder learning (Paas, Renkl \& Sweller, 2003). In the
context of collaborative learning, we understand the cognitive load induced by
mutual modeling as part of Schwartz’s (1995) notion of effort towards a shared
understanding. For instance, CSCL researchers expanded the use of 'collaboration
scripts' (Aronson,?). A script is a pedagogical method that collaborative
learning activities in order to foster the emergence of productive interactions
such as argumentation, explanation or conflict. Conflict-resolution scripts such
as the ArgueGraph (Dillenbourg \& Hong, 2008) forms pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, JIGSAW
scripts provide peers with different but complementary knowledge for augmenting
(reasonably) the effort group members have to engage to reach a shared solution.

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
%people, A and B, manipulate objects. Then A leaves and B hides away one
%object. When A comes back, we ask the child ``where do you think A will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world (A still thinks the object is at its original
%position since he did not see B hiding it).

\section{Mutual Modeling}

In order to build a shared understanding, do partners have to build a
representation of each other's knowledge? We refer to \emph{mutual modeling} as
the process of inferring one's partner mental states. Any claim that students
carry out a detailed monitoring of their peers would be as incorrect as any
claim that they do not maintain any representation at all. If mutual modeling
had to be permanently detailed and accurate, subjects would obviously face a
huge cognitive load. Conversely, peers could not collaborate without some
minimal amount of mutual modeling. For instance, A cannot disagree with B
without knowing that B has a different opinion. The mutual model can be implicit
(A is not aware of what he knows about B), by-default (I believe that B beliefs
what I believe unless contrary evidence), opportunistic (A does not model B
unless the conversation requires it), global (A infers B's beliefs based on
categories such as age, culture or profession) and, of course, it can be
incorrect… but it can not remain empty. Dialogues include many instances of
utterances such as "I thought he would do that" (first level of mutual
modelling) or even "He thought I would do that but I intended something else."
(second level of mutual modelling).

The content of mutual models ranges from 'dispositional' versus 'situational'
aspects. The 'dispositional' aspects refer to A's representation of B's long
term knowledge, skills or traits. It is thus closely related to the notion of
transactive memory~\cite{wegner1987transactive, moreland1999transactive}.
'Situational' aspects refer to A's representation of B's knowledge, behavior or
intentions specifically activated in the situation in which A and B are
collaborating, some of them being valid for 2 seconds, other ones for 2 hours.
Here are examples of fragments that constitute A's model of B regarding to
aspects X, noted $Model(A,B,X)$:

\begin{itemize}

    \item $Model(A,B, knowledge)$: What does A know about B's knowledge with
        respect to the task at hand or, inversely, about B's knowledge gaps?
        When can A consider B's statements as reliable?

    \item $Model(A,B, skills)$: What does A know about B's skills with respect to
        the task at hand? May A expect B to perform well in a specific subtask?
        The effectiveness of division of labor depends on the quality of this
        mutual model.

    \item $Model(A,B, goals)$: What does A know about B's intentions with respect
        to the project, including B' motivation and commitment? Can A trust B
        when B promises to deliver?

    \item $Model(A,B, task)$: What does A know about B's representation of the
        situation and the task: does A knows whether B has the same
        understanding of the problem at stake?

    \item $Model(A,B, plans)$: What does A know about B's strategy. Does A
        understand why B did what he did? Is A able to anticipate what B will do
        next?

    \item $Model(A,B, "urgent")$: What does A about know B's understanding of A's
        last utterance: does ‘urgent’ means now, ASAP and ‘not too lat’ ?

    \item We could continue the list of what is X is \model{A}{B}{X}: beliefs,
        emotions, history, status, …

\end{itemize}

We have different levels of mutual modelling. If A states "B thinks I am god in maths", we
are the second level of mutual modelling:  $Model(A, B, Model(B,A, knowledge="good in
math"))$. There is an infinite regress of nested models: If A says "B knows that
I don't expect him to solve this statistics problem" corresponds to $Model(A, B,
Model (B, A (Model (A,B, statistic-skills)))$.

A partner model is probably not a "box", \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
appearance. Scholars studied how initial modeling impacts communication. In
their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
Ginsburg~\cite{slugoski1993attribution} pretended to their subjects that their
(fake) partner had or had not received the same information. They observed that
the subjects adapted their dialogue by focusing the explanation on the items
that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
different initial strategies in forming queries depending on who they were told
their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which A and B attended together~\cite{clark2002definite} in the physical space
or in our cultural space (\eg ‘09-11’). Co-presence means that we can refer to
a shared objects and event,-s it does of course not imply that we give them the
same meaning. Namely, a shared screen does not mean a shared
understanding~\cite{dillenbourg2006sharing}.

After initialization, mutual models are updated along the collaborative work
through verbal and non-verbal interactions. A default inference rule is that "my
partner agrees with me unless he disagrees", which reject the critiques that
mutual modeling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing one's partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle~\cite{roschelle1995construction} reformulate
the design of CSCL interfaces as providing ways for peer to detect and repairs
their misunderstanding. In CSCW, researchers have explored various so-called
"awareness tools", \ie functionalities that inform A is about B's actions that A
could not directly perceived because B was working on a different subset of the
virtual space. Different awareness tools will be addressed in this contribution
since they allowed us to manipulate experimentally the mutual modelling activity.

%\subsection{State of the Art}
%
%\paragraph{Building a model of a human during the interaction} Perspective
%Taking is a human ability which allows one to put him/herself in another
%person's point of view. Studied in psychology
%literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
%interacting with people by allowing one to reason on others' understanding of
%the world in terms of visual perception, spatial descriptions, affordances and
%beliefs, etc.  In the last years these notions have been gradually employed in
%HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
%information about a teacher's visual perspective in order to learn a task.
%~\cite{Johnson2005} apply visual perspective taking for action recognition
%between two robots.~\cite{Trafton2005} use both visual and spatial perspective
%taking to find out the referent indicated by a human partner.
%
%The cognitive model that the robot builds for the agent it interacts with is
%still simple and mostly focused on geometric features (who sees what? What are
%our relative positions? etc.). Extending this knowledge with more subtle
%perceptions (emotional state for instance) remains to be done.
%
%\paragraph{Joint action}
%Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
%emphasize that collaborative tasks have specific requirements compared to
%individual ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realize it, they must show their commitment to
%the goal during execution, etc. Several robotic systems have already been built
%based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
%have shown benefits of this approach. They have also shown how difficult it is
%to manage turn-taking between communication partners and to interleave task
%realization and communication in a generic way. Finally, today only few
%systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
%levels.
%
\section{A model for mutual modelling evaluation}

To report experiments on mutual modeling, we suggest
the notation \model{A}{B}{X} to denote ``A knows that B knows X". This is a
simple but useful reduction to reason on mutual modelling. This
notation does not mean A has an explicit, monolithic representation of B: it
must be understood as an abstraction referring to complex socio-cognitive
processes. Besides, we refer to the \emph{degree of accuracy} of the model as
\Model{A}{B}{X}.

We parametrize and assess the mutual modelling effort through 3 variables:

\begin{enumerate}

    \item Tasks vary a lot with respect to how much they require.  The grounding
        criterion -- denoted \groundingcriterion -- refers to how
        important it is to mutually share a piece of information X to succeed
        the task T. It can be computed as the probability to succeed T despite
        the fact X is not grounded. $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be
        estimated from the correlation between \Model{A}{B}{X} and task
        performance.

    \item Before any specific grounding act, there is usually a non-null probability
        that X is mutually understood by A and B: because X is part of A's and
        B's cultures, because it is manifest to co-present subjects or simply
        because there is not much space for misunderstanding or disagreement
        about X. We could not simply collaborate without a high level of initial
        grounds. We note the theoretical accuracy of initial grounds:
        $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.

    \item The cost of grounding X refers to the physical and cognitive effort
        necessary to perform a grounding act $\alpha$: a verbal repair (\eg
        rephrasing), a deictic gesture, a physical move to adopt one partner's
        viewpoint, etc. This cost varies according to media features ( Clark \&
        Brennan (REF)).

\end{enumerate}

Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
content X at time $t_1$ during task T in order to increase \Model{A}{B}{X}
at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
costs (3)~\cite{traum1996miscommunication}:

\begin{equation} \label{eq:probrepair}
p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \frac{(M^{\circ}_{min}(A,B,X,T) -
M^{\circ}_{t_1}(A,B,X))}{cost (\alpha)}
\end{equation}

This is qualitative summary more that a real equation since several parameters
are hard to quantify (\eg the cost an communication acts depends upon the user
as well) but it clarifies the parameters of the experiments we propose hereafter.

\begin{itemize}
    \item \Model{A}{B}{X}: Our experiments address different contents that can be
        represented in mutual models:

    \begin{enumerate}
        \item \model{A}{B}{emotion}: how accurately A perceives B's emotional state
            (study 3);

        \item \model{A}{B}{actions} is about how well A guesses what action B has
            done (study 2) or will do next  (study 1)

        \item \model{A}{B}{knowledge}: how accurately A estimates B's knowledge
            with respect to the material they learn together (study 4 and 5)

    \end{enumerate}

    \item \Model{A,B}{X}{T} Our studies concern various collaborative tasks:
        argumentation (study 3), games (study 1 and 2) and concept mapping
        (study 4 and 5). By varying the tasks, we do actually vary the grounding
        criterion but they all require a "reasonably high" grounding criterion.
        All tasks are about building a solution or a representation together; we
        did not address everyday conversations.

    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$ Along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university: they hence had some general common
        grounds.  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a JIGSAW script.

    \item $cost(\alpha)$ In all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modeling emotions is "cheaper" with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provide A with information about
        what B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
        and, like any prosthesis, they many augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

\end{itemize}

Each of the studies that we present in the next section have been design in such a way that, while tractable on a
robot, mutual modelling plays a key role for the task performance: the tasks
require a ``reasonably high" grounding criterion \groundingcriterion,
and all tasks are about building a solution or a representation together.

Along the same reasoning, the initial degree of common grounds \inigrounding
should be rather low (and hence the difference between initial and required
degrees rather high) in order to make mutual modelling effort more observable.
While we can exactly control the actual initial knowledge available to the
robot, estimating \inigrounding still requires special care during the design of
the experiments: as we have previously shown~\cite{lemaignan2014dynamics,
lemaignan2014cognitive} and because the interaction takes place between a human and
a robot, the initial human model of the robot can widely vary depending on
several factors like the user familiarity with technology, the effect of
novelty, the user expectations or the anthropomorphic features of the robot.

More generally, while we have explicit access to \model{R}{H}{X} (from the facts
stored in the robot belief state, as illustrated in
Table~\ref{table|beliefsfig7}), estimating \Model{R}{H}{X} also requires
\model{H}{H}{X}, which depends on the meta-cognitive competencies of the human
partner.

Conversely, measuring \Model{A}{B}{X} is methodologically difficult. We
discriminate two steps, first to capture \model{A}{B}{X} and then to
estimate \Model{A}{B}{X}.

\begin{itemize}
    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what (s)he
        believes about what $B$ knows, feels, intends to do, etc. This raises
        obvious methodological concerns since such a question triggers a
        modeling process beyond what it would 'naturally' be. To avoid this
        bias, one can estimate mutual modeling after task completion. Then, the
        obvious drawback are memory losses and by post-hoc reconstruction.
        The first method was used in study 2 and the
        second one in the other studies. Another option would be to rely on external behavioural metrics like
        eye-tracking: we hypothesise for instance that the fixation time
        reflects the efforts engaged by the human to understand, hence, model,
        the others. Such an approach has however not yet been investigated in the presented studies.

    \item {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
        need to access the reference model \refmodel{B}{X} to estimate its
        \emph{accuracy}. Since we can only indirectly access it via what $B$
        reports (\ie \model{B}{B}{X}), accuracy can be estimated in 2 ways:

        \begin{itemize}

            \item Subjective accuracy: In study 3, we compute
                \Model{A}{B}{X} by measuring if $A$ describes
                $B$'s emotions in the same way $B$ reports its emotions 
                ($\M{A}{B}{X} = \M{B}{B}{X}$).

            \item Objective accuracy: In studies 4 and 5, we compute
                \Model{A}{B}{X} by comparing \Model{A}{B}{K} to $B$'s
                actual knowledge as it has measured by a test.

        \end{itemize}

\end{itemize}

\subsubsection{Studies}

The experiments we report here address mutual modelling across different tasks, some with
dyads, other with triads. They were conducted along 6 years in two different
institutions by different researchers. They used different independent,
intermediate and dependent variables. Nonetheless, we were able to address a few
questions across these studies. We start with 3 simple hypotheses about
$\mathcal{M}^{\circ}(A,B)$:

\begin{itemize}
    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon A's ability or effort
        to model B,

    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  B's ability of
        effort to help A to model him/herself

    \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
        interactions among A and B

\end{itemize}

$\mathcal{H}_{2}$ is also called a second level modeling hypothesis since H
needs to monitor B to see if A understood him/her
$\mathcal{M}(B,(A,\mathcal{M}(A,B)))$

Our main issue is the symmetry questions: what is the relationship
(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and
\Model{B}{A}{X}? A low symmetry would mean that mutual modeling is
mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
might support $\mathcal{H}_{2}$ since there is a low probability that randomly
formed pairs integrate peers with the same level of mutual modelling skills.

\begin{figure}[htb]
\centering

\begin{tikzpicture}[scale=0.5]

\draw(0,0) node[anchor=north] (A) {\scriptsize A};
\draw(10,0) node[anchor=north] (B) {\scriptsize B};
\draw(5,2) node[anchor=north] {\scriptsize \Model{A}{ B}{ X}};
\draw(5,-2) node[anchor=north] {\scriptsize \Model{B}{ A}{ X}};
\draw[dashed, <->] (5,1) -- (5,-1.9) node[midway, sloped, above] {$\Delta_1$};
\draw[->] (A) to[bend left] (B);
\draw[->] (B) to[bend left] (A);

\end{tikzpicture}

\caption{The mutual modelling symmetry question $\Delta_1 =  \Delta(\mathcal{M}^{\circ} (A,B,X),
\mathcal{M}^{\circ} (B,A,X))$}

\label{mm_symmetry}
\end{figure}

With triads, we may compute the accuracy of 6 models:
\Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
\Model{C}{B}{X} and \Model{B}{C}{X}. This enables two
triangle questions (see Fig.~\ref{mm_triangles}):

Do A and B have the same accuracy when modeling C? $\Delta_2 =
\Delta(\M{A}{C}{X}, \M{B}{C}{X})$ If it is the same,
$\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain over $\mathcal{H}_{1}$

Does C model more accurately A than B? $\Delta_3= \Delta(\M{C}{A}{X},
\M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.

\begin{figure}[htb]
\centering
\subcaptionbox{}{ 
    \begin{tikzpicture}[scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize \Model{A}{ C}{ X}};
    \draw(5,-2) node {\scriptsize \Model{B}{ C}{ X}};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above] {$\Delta_2$};
    \draw[->] (A) to (C);
    \draw[->] (B) to (C);

    \end{tikzpicture}
}
\subcaptionbox{}{ 
    \begin{tikzpicture}[scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize \Model{C}{ A}{ X}};
    \draw(5,-2) node {\scriptsize \Model{C}{ B}{ X}};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above]
    {$\Delta_3$};
    \draw[<-] (A) to (C);
    \draw[<-] (B) to (C);

    \end{tikzpicture}
}
\caption{The mutual modelling triangle questions}

\label{mm_triangles}
\end{figure}

In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
whether the accuracy of mutual modelling depends more upon the modeller's effort
or the modellee's behaviour.

\paragraph{The rectangle questions}

We could go further by comparing self- versus other modeling $\Delta_4$ in
Fig.~\ref{mm_rectangle}) as an indication of metacognitive skills. We could also
see if modeling skills depend upon what aspects are being modeled (X or Y),
which would explain vertical differences ($\Delta_5$ in
Fig.~\ref{mm_rectangle}). We do not further develop these questions because we
don’t have the necessary data in our studies.

\begin{figure}[htb]
\centering

\begin{tikzpicture}[scale=0.5]

    \draw(0,0) node (a) {\scriptsize \Model{A}{ B}{ X}};
    \draw(10,0) node (b) {\scriptsize \Model{A}{ A}{ X}};
    \draw(10,-4) node (c) {\scriptsize \Model{A}{ A}{ Y}};
    \draw(0,-4) node (d) {\scriptsize \Model{A}{ B}{ Y}};
    \draw[<->] (a) -- (b) node[midway, below] {$\Delta_4$};
    \draw[<->] (b) -- (c);
    \draw[<->] (c) -- (d);
    \draw[<->] (d) -- (a) node[midway, sloped, above] {$\Delta_5$};

\end{tikzpicture}

\caption{The rectangle questions}

\label{mm_rectangle}
\end{figure}

This simple notation does not pretend to provide a mathematical account of
mutual modeling but to be more systematic in describing the following
experiments.

\section{Studies}

\subsection{Study 1: Effect of an awareness tool on \Model{A}{B}{x} in a virtual
game}

\paragraph{Questions}

We studied the impact of an awareness tool on group performance and mutual
modeling. (Nova, Wehrle, Goslin, Bourquin \& Dillenbourg (2006). The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: "I did that because I thought that
you would do that". Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

\paragraph{Experimental setting}

SpaceMiners is a 3D game that involves two players collecting minerals located
in asteroids. To do so, they shoot drones through the space after choosing their
initial direction and speed. Once launched, the trajectory of drones is only
influenced by the gravity of planets and by trajectory modification tools.
During the experiment, the teams were confronted with three increasingly complex
situations. The experiment was 2 hours long: a 30 minutes tutorial and 3 levels
of 30 minutes. Players uses using a regular joystick and communicated with each
other through an audio channel.

The independent variable was the availability of an awareness tool that shows to
player A  the location and gaze direction of player B. In the awareness
condition, players could switch to the 'scout mode' where they could view what
their partner was looking at. We hypothesize that this would enable subjects to
more accurately infer their teammate's intentions. Each player sat in front of a
distinct computer located in different rooms.

\paragraph{Subjects}

Thirty-six persons participated in this study, all native French speakers. We
constituted 18 pairs of participants (N = 18) who were not familiar with each
other. The pairs were assigned randomly to either the control condition (without
the awareness tool) or the awareness condition (with the awareness tool).

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image4.png}
        \caption{Snapshot of the SpaceMiner Game}
        \label{spaceminer}
\end{figure*}

\paragraph{Variables}

Task performance was measured by the score reached by the two subjects after
three situations. The effort of mutual modeling was measured as the ratio of
time that players would spend in the scout mode (divided by total time), which
is the time during which players are not performing their own actions but
monitoring their partner's actions.

In order to evaluate m°(A,B) during the task, we used two questionnaires (Figure
4) that were displayed during each of the three games, as a transparent layer
appearing over the game display. The first questionnaire concerned the player’s
intended actions. The second questionnaire asked each player about what he
thought her or his partner was intending to do. Some answers were identical in
both questionnaires (like “adjusting a shot”) while others were reversed ("to
guide my partner" versus "to guide me"). We chose this subjective measure of
accuracy (∆ M(A,B), M(B,B)) rather than an objective measure (i.e. model (A,B)
is compared to B's next action) because some the activities proposed by the
questionnaire were not observable by the environment (e.g. establishing a
strategy). We counted M°(A,B,activity,gamei) as the number of common answers
between questionnaires M(A,B) and M(B,B) in each game and computed the average
value a across the 3 games.

\begin{figure}[ht!]
        \centering
        \includegraphics[width=\columnwidth]{image5.png}
        \caption{M(A,A) and M(A,B) questionnaires in SpaceMiners (translated
        from French)}
        \label{study1:questionnaires}
\end{figure}

\paragraph{Results}

Grounding criterion. The grounding criterion was high: the correlation between
M°(A,B) and task performance was 0.42, p = 0.05. Pairs with an accurate mutual
model reached higher scores. A regression analysis confirmed the positive and
significant relation between group performance and mutual modelling accuracy (β
=54, p = .02).

Study-specific questions. The awareness tool permitted higher group performance,
but it did not improve the accuracy of the mutual model. Since teams were free
to use the awareness tool or not (the 'scout' mode), we performed a post-hoc
split of players depending on how much time they used it. The split point was
the mean of time spent in the scout mode and it led to the constitution of two
groups made up of 12 individuals “short time in scout mode” and 24 individuals
“long time in scout mode”. A two-way analysis of variance conducted on these
contrasted groups revealed that pairs in the awareness condition who spent more
time in the scout mode reached higher levels of M°(A,B)(F = 8.02, p = 0.015). Of
course, a post-hoc split does not support a causal direction. An alternative
explanation could be that good modellers are more social and hence appreciate
the awareness tool.

Symmetry question. We computed intra-class correlation as described by Kenny et
al. (1998) from the answers to the cross-questionnaires. Considering all pairs
in both conditions, we found a positive and significant correlation (r = .38, p
< .05) between M°(A,B) and M°(B,A). Interestingly, this was higher in the
control group (r = 0.44) than in the experimental group (r = 0.24). Actually, ∆
(M°(A,B),M°(B,A)), i.e. the absolute differences between the models accuracy,
was not significantly different with or without the awareness tool ( F [1,13]=
0.144, p > 0.5). This result could be explained by the fact that the players
without awareness tools communicated more.

We cannot answer the triangle questions since this study was conducted with
pairs-

\paragraph{Discussion}

How to interpret a correlation of 0.38 between M°(A,B) and M°(B,A)? If it was
null, modelling would be interpreted as an individual process that varies
according to personality traits or cognitive skills: some persons tend to model
more accurately their peer while others don't care or can simply not. If the
correlation was close to 1, we could conclude that the modelling accuracy is an
emergent property of teams since there would be low probability that, by chance,
low accurate modellers are paired with accurate modellers and conversely. Our
correlation is between these two extremes but the very notion of significant
intra-class correlation tend to indicate that one may not consider peers as
independent subjects, which is an argument for the 'interaction' hypothesis
(H3).

\subsection{Study 2: Effect of an awareness tool on M°(A,B)  in a pervasive
game}

\paragraph{Questions}

This study concerns a collaborative game that occurred in the physical space
(Nova et al , 200?). We studied whether players build an accurate model of the
path followed by their partners because this path reflected their problem
solving strategy. We used an objective measure of M°(A,B): the distance between
where A believes B has been walking and where B actually went. The main
hypothesis concerned the effect of awareness tools on group performance and on
M°(A,B).

\paragraph{Experimental setting}

CatchBob is a mobile game in which groups of 3 players have to solve a joint
task. The game was played on EPFL campus. Participants had to find a virtual
object ('Bob') and "to catch it" by forming a triangle around it. Players used a
Tablet PC that displayed a map of the campus and an indication of their personal
distance from Bob. Their annotations on the map were shared with the two other
players (A could see what B and C wrote). These annotations faded out after a
few minutes to avoid covering the full display. The awareness tool displayed the
location of the two other players on the map. The availability of location
awareness tool constituted our independent variable.

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image6.png}
            \caption{A's drawing A of B's path.}
        \end{subfigure}
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image7.png}
            \caption{Actual path followed by B.}
        \end{subfigure}
        \caption{Bla bla}
        \label{study2:paths}
\end{figure*}

\paragraph{Subjects}

Ninety students participated in this experiment. We chosen only students from
our University because knowing the campus geography had an impact both on group
performance but also on mutual modeling: it is hard to represent the path of
someone across some space without having a mental map of this space. We selected
groups of students who knew each other. We assigned 10 triads to each of our
three experimental conditions. Each condition was made up of approximately 25%
of women, but we did not control gender repartition within each triad.

\paragraph{Variables}

The independent variable was the awareness tool: the control condition (without
tool) and two experimental conditions: synchronous awareness (display current
position of each player) and asynchronous awareness (display current position of
each player and their recent spatial trace).  As a dependent variable, we had
the task performance which was the distance covered by the team to catch Bob and
M°(A,B). To estimate M°(A,B), we asked players to draw on paper their own path
and the path of each of their partners after the game. This enabled us to
calculate the number of errors players made while drawing the path of their
partners. We compared the path that player A attributed to B with B's real paths
recorded by the system and the same for A\&C or B\&C as depicted on Figure 5.

M°(A,B) is the sum of errors made by A about B’s paths. An error was either
drawing a place where the partner had not been or not drawing a place where
he/she had gone. One could argue that M°(A,B) is biased by the subjects' ability
to translate their trajectories souvenir into a map drawing. However, 85% of
subjects made no mistake at all when drawing their own path. We therefore
consider mistakes in their partners' path as being due to a lack of mutual
modeling accuracy instead of being due to spatial reasoning skills.

\paragraph{Results}

Grounding criterion. The correlation between M°(A,B) and the task performance
(path lengths) was low: 0.15. Using a post-hoc split on M°(A,B), there is no
significant difference between the performance of the groups with high and low
M°(A,B)  (F = 1.45, p = 0.24). Conversely, a post-hoc split of the groups
according to their performance did not show any significant differences on
M°(A,B) (F = 1.16, p = 0.29).

Study-specific questions. There was no significant difference regarding the task
performance. However, our surprise was that the absence of the awareness tool
was related to a higher M°(A,B): players better remembered their partners’ path
if they did not see their position permanently! We will not enter into the
details of these results but, to be short, teams without awareness tool made
more annotations on the map while permanent MLA has an underwhelming effect
(Nova et al., 2005).

Symmetry question. The correlation between M°(A,B)  and M°(B,A)  is positive (r
= .41) and significant (p < .01);  the more A makes errors about B, the more B
does the same.

Triangle questions. About ∆2, the correlation between M°(A,B) and M°(A,C)is
significant: r=0.30, p <.01. About ∆3, the correlation between M°(A,C) and
M°(B,C) is significant: r=0.43, p <.001.

\paragraph{Discussion}

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults (Liben et al, 1981), a significant correlation could indicate that mutual
modeling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modeling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modeling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one’s partner path.

\subsection{Study 3:  Effect of media richness on M°(A,B) in argumentation}

\paragraph{Questions}

The aim of this unpublished study was to evaluate the effect of media richness
on M°(A,B,emotions). The hypothesis was that video communication would lead to a
better M°(A,B) than audio only since emotions often impact on facial
expressions.

\begin{figure*}[ht!]
        \centering
        \includegraphics[width=0.9\textwidth]{image8.jpg}
        \caption{Example of argumentation graph}
        \label{study3:argumentation_graph}
\end{figure*}

\paragraph{Experimental settings}

Triads had to address an emotional societal debate: authorizing or not adoption
by homosexual couples). They worked on-line and had to structure their
argumentation with a shared concept map tool (TeamWave) as illustrated in Fig.
6. Ten groups had only an audio connection while ten groups had audio plus
video. The video communication was provided by a web cam and the software
IVisit. For the audio link, we used microphones, headsets and a software called
BattleCom. In the audio+video condition, the screen was divided in three
sections. The main part was devoted to the concept map window (24 X 24 cm). In
the remaining part appeared the image of the two peers (8.5 X 8.5 cm each). In
the audio condition, this video zone was left empty so that the size of the
concept map was equal in each condition. The subjects were located in the same
room, separated by mobile walls. Despite their headsets, it is not impossible
that they could hear non-verbal audio cues (e.g. tapping the sole with feet).
The task last in average 61 minutes.

\paragraph{Subjects}

Twenty triads from the University of Geneva participated to this experiment. We
formed groups of subjects who knew each other, i.e. they had ab initio mutual
models: the task required the discussion of sensitive issues which required to
feel quite comfortable with peers. There were 36 women and 24 men, but, since
groups were formed a priori, we could not control the balance of men and women
in each condition.

\paragraph{Variables}

The independent variable was the presence of not of video link. The dependent
variable M°(A,B) was subjectively by using 3 questionnaires: in the first one, A
described his or her own emotions M(A,A) , while in the two other questionnaire,
A describes B’s and C’s emotions.   The questionnaire included 18 items (7 steps
Likert Scale) describing emotions labeled as adjectives: anxious, enthusiastic,
agitated, proud, excited, quiet, calm, stressed, bored, upset, relaxed,
irritated, determined, hostile, active, etc. This list of emotions was
elaborated from (REF?). The proposed emotions were . M(A,B) is a vector of 18
numerical values {x1 … xi} corresponding to their answers on each questionnaire.
M°(A,B) has been computed as the correlation between the two vectors :

M°(A,B,emotions) = ∆M(A,B,{x1 … xi}, M(B,B,{y1, … ,yi}) = r ({x1 … xi}, {y1, …
,yi}))

\paragraph{Results}

\begin{figure*}[ht!]
        \centering
        \includegraphics[width=0.9\textwidth]{image9.pdf}
        \caption{Average values of M°(A,B,X) where X is one of the proposed
        emotions (max = 7).}
        \label{study3:deg_m_values}
\end{figure*}

Grounding criterion. The maps produced by teams were ranked. We use the average
rank as a estimation of the team performance and correlated with the average of
the 6 values of M°(A,B) per team (M°(A,B), M°(A,C), M°(BA),…): The correlation
is -0.22: teams with a good M°(A,B) team to be better ranked.

Study-specific questions. Our hypothesis about media richness is rejected: the
average M°(A,B, emotions) was .62 (SD .28) with the audio+video condition and
.71 (SD .19) in the audio condition (F=3.4, p: 0.065). Our interpretation is
that viewing what one's partner sees (shared graphical editor) was, in this
task, more important than seeing each other (Gaver et al 1993; Anderson et al.,
2000).

Symmetry questions. Regarding the main items, the mutual modeling of emotions,
the correlation is .25 or .36  a reprendre avec - Patrick

Triangle questions. Check (∆2) . For (∆3) If we compute for each subject A, the
correlation between the vectors of emotions in M(B,A) and the vector in M(C,A),
and take the average across all 60 subjects, we obtain a very impressive
correlation of .66. This average is about the same in each condition.

Rectangle questions We cannot address the relationship (∆4) between self-and
social accuracy here because we don't have an estimation of self-accuracy:
subjects describe their emotions but we have no way to check if there are
correct. However, by measuring M°(A,B) on  18 emotional labels, we can however
have a glimpse about ∆5: how M°(A,B,X) varies according to X. Figure 7 shows the
range of modelling errors: the difference between M(A,B) and M(B,A), on a scale
of 7, is in average of 0.3 for the emotion "discasted" up to 1.9 for the emotion
"calm". This is probably specific to the variety of scales  (SD= 0-5 for
"discated" versus .7 for "calm"), our point is not to interpret this too far,
but to show that there are large varations even with one area  (perceiving
emotions).

\paragraph{Discussion}

Chek si il faut modifier les resultats de la symmetries

\subsection{Study 4:  Effects of a script on M°(A,B)  in concept mapping}

\paragraph{Questions}

This study investigated the effect of a collaboration script on collaborative
learning (REF).. The script chosen is a JIGSAW: two students receive different
but complementary subsets of the knowledge (texts) which have to be integrated
to build a shared concept map.  This script increases the cognitive effort to
build the map, not only to recon ciliate the viewpoints of each team members
but, before that, to find out what the other knows.

\paragraph{Experimental settings}

The instructional material consisted of an explanatory text about the
neurophysiologic phenomenon of “action potential”. The text was divided into 3
chapters. The peers were located in two rooms equipped with the same computer.
The experimental session lasted around 90 minutes and consisted of 6 phases:
Participants used two software components, CmapTools (http://cmap.ihmc.us/) and
TeamSpeak (http://www.goteamspeak.com/).

1. As a pre-test, participants were asked to write down all they knew about the
neuron and its functioning (5 minutes).

2. Participants were instructed to read a text (12 minutes).

3. Participants were asked to build individually a concept map in order to
graphically represent what they learnt from the text (10 minutes).

4. Dyads had to built a concept map during 20 minutes, communicating by audio.
The screen layout was structured into three areas (see Figure 8)

5. Participants were invited to individually complete a knowledge test (15-20
minutes).

6. Participants where asked to estimate their own- and their partner’s final
knowledge in a questionnaire.

Figure ?...

\begin{figure*}
    \centering
    \begin{tikzpicture}[scale=0.8]
        \node[anchor=south west,inner sep=0] at (0,0)
            {\includegraphics[width=0.65\textwidth]{image10.png}};

            \node[rectangle,draw, thick] at (11,6.2) {Self map};
            \node[rectangle,draw, thick] at (11,3) {Partner map};
            \node[rectangle,draw, thick] at (3,2) {Collaborative map};
    \end{tikzpicture}
    \caption{The group concept map and the individual concept maps in study
    4}
    \label{study4:concept_map}
\end{figure*}

\paragraph{Subjects}

Fifty-eight first year students from EPFL, 47 men and 11 women, with a mean age
of 20.46 were remunerated for participation. Dyads were randomly assigned to one
of the two experimental conditions. The men/women ratio was equivalent in each
condition. Participants did not know each other before the experiment. Students
from Life Sciences were not recruited since they had too high background
knowledge on the learning domain.

\paragraph{Variables}

The independent variable, script versus no-script, was implemented by the
difference of texts that individuals had to read. In the “same information” (SI)
condition, the same text was given to both partners. In the “complementary
information” (CI) condition. it was divided into two sub-texts, one about the
electrical processes of the neuron while the second one about the chemical
processes. These two versions were equivalent in terms of number of information
pieces.

The dependent variables were the post-test scores, but this paper is mostly
concerned by the M(A,B, knowledge).  In phase 6, participants were ask to
estimate (7-point Likert scale) their own- and their partner’s outcome knowledge
with respect to each chapter of the zexz. The order of questions about oneself
and about the other was counterbalanced across participants. We measure M°(A,B)
in an objective (1 \& 2) as well as in a subjective way (3).

(1) M°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – score (B, chapteri) |

(2) M°(A,A, knowledge) = Σi=1to3  | M(A,A, chapteri) – score (A, chapteri) |

(3) m°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – M(B,A, chapteri) |

\paragraph{Results}

Grounding criterion. In this task, the grounding criterion was low, but we did
not evaluate task performance (e.g. the quality of the jointly produced concept
map) but the learning gains. The correlation between M°(A,B) and A's learning
gains is not significant (β =.08, ,NS, N = 60). It is also not significant
within each condition.

Study-specific questions. We performed a non-parametric Mann-Whitney test on
post-test scores for electric questions and a one-way ANOVA on scores for ionic
questions (Levene tests for homogeneity of variances: p = .02 and ns,
respectively. Results  did not show any significant difference between the “same
information” (SI) condition and the “complementary information” (CI) condition,
neither for electric questions [U = 388.50, z = -0.88, ns] nor for ionic
questions [F(1, 58) = 0.17, ns].  The effect of scripts on M°(A,B) was not
significant (F(1, 58) = 0.78, ns) when considering the absolute difference
between M°(A,B) and B's post-test score. However, A tended underestimate B's
score in the SI condition (M = -2.06) and to overestimate it in the CI condition
(M = 1.21). (F(1, 58) = 6.44, p<.01). Regarding to M°(A,A), there was no
significant difference between conditions.

Symmetry question. The interclass correlation between M°(A,B) and M° (B,A) is
not significant, but almost (r=.26 (F(1,29)=1.71, p= 0.075)). Actually, it is
significant when students read the same text (SI condition: r= 0.43,
F(1,15)=2.53, p<.05) but not when they read different texts (r=.13, F(1,12)=1.3,
NS).

Triangle questions do not apply here since we worked with dyads.

Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is globally not
significant (r=0.05) as well as not significant in each condition: someone good
at self-modeling is not necessarily good at modeling someone else and
vice-versa.  Discussion

In this experiment, the symmetry of mutual modelling is only found when subjects
are in conditions of symmetry of information. This consolidates the hypothesis
of M°(A,B) being a property of interactions in context rather than A's cognitive
attitude.

\subsection{Study 5:  M°(A,B, knowledge)  in concept mapping}

\paragraph{Questions}

This study investigates if M°(A,B, knowledge) is related to learning outcomes by
comparing teams with or without a knowledge awareness tool (KAT), i.e. a tool
that informs A about B's knowledge as measured through a pre-test.

\paragraph{Experimental setting}

The peers were located into two different rooms. A complete description of the
study is provided in Sangin et al (2008).The experiment lasted 90 minutes. It
started with the same two steps as in study 4.

3. Subjects passed a pre-test, with ten questions per chapter.

4. Participants had 20 minutes to draw a collaborative concept map reporting the
content of the texts. They were able to communicate orally through headsets.  We
used Tobbii eye tracking devices to record their gazes.

5. The post-test included the same items than the pre-test but in a different
order.

6. Finally, participants were asked to estimate their partner’s knowledge at the
post-test for each of the three chapters on a 7-point likert-like survey.

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image11.png}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

\paragraph{Subjects}

Sixty-four first year EPFL students (18 women, mean age = 21.2 years)
participated to the study. They were involved in curricula that do not imply
advanced neurophysiologic notions. They were randomly assigned to conditions and
did not know each other before.

\paragraph{Variables}

The participants of the experimental condition group were provided with the KAT
on the bottom part of the screen (see Figure 9): each line represents the score
obtained by the partner at the pre-test for a chapter. Participants did not see
they own score, but they usually started their discussion by exchanging this
information.

\paragraph{Results}

Grounding criterion. A linear regression revealed a positive relation between M°
(A,B) and the collaborative learning gain of the pair (AB): β = .401, p < .001,
r2adj. =.15, large effect. This relationship was not significant in the previous
experiment which as been conducted with the same task: using the same task
should produce the same grounding criterion This probably due to the fact that
M° (A,B) was influence by the KAT.

Study-specific questions. The t-test reported a significant difference between
the KAT condition participants (M = 13.4\%) and the control group [M = 3.6%, t(1,
60) = 2.73, p < .01,Cohen’s d = 0.7, medium to large effect]. Providing learners
with cues about the prior-knowledge of their partner enhances their
collaborative learning. As a treatment check, we found a positive and
significant correlation between the amount of gazes on KAT (using eye tracking
devices) and the learning gains (r(22) = .54, p = .01). A detailed analysis
revealed that subjects look at KAT to assess their peer's credibility when (s)he
provided new information.

The KAT has a significant effect on M°(A,B): peers more accurately estimated
their partners knowledge (M = 1.11) than those in the control condition [M =
.98, t(1, 60) = 3.19, p < .01, Cohen’s d = 0.83, large effect]. This is a
trivial result since the KAT provided them with an initial M(A,B). However,
subjects have to predict the post-test score while the KAT informed them about
the pre-test score. Actually, pairs in the KAT condition produced significantly
more instances of 3 specific categories of interactions: (1) elaborated
utterances with rich contents,  (2) utterances asking about the other's
knowledge such as "Did you understand how transmission works?" (3) utterances
describing one's own knowledge (M(A,A)) such as "I don’t remember the Ranvier’s
thing…" . These 3 categories provide different account of M°(A,B): (1) as an
effect of the quality of interaction, (2) as an effect of A's effort to model B
or (3) as B's effort to give cues to A about his own knowledge.

We examined M° (A,B) as potentially mediating the effect of the KAT factor on
the relative learning gain. A linear regression confirmed that M° (A,B) was
significantly related to the KAT-factor (β = .381, p < .01, r2 = .15). The
KAT-factor was also significantly and positively related to the RLG (β = .332, p
< .01, r2 = .11). We then tested the relation between the independent variable
(KAT) and the dependent variable (gains) when controlling for the mediating
variable (M° (A,B)). A multiple regression showed that the KAT-factor was no
longer a significant predictor (β = .210; p = ns) whereas the M° (A,B) was still
a significant predictor (β = .32, p < .01). Thus, it can be concluded that M°
(A,B) mediated the KAT-factor’s effect on the learners’ RLG. The Sobel (1982)
significance test for indirect effects was significant [z = 1.99, p < .05].

Symmetry question.  We did not find an intra-pair correlation (ICC = 0.05, NS).
Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is \fixme{???
MIRWEIS A FAIRE}.

\paragraph{Discussion}

Despite the fact that the learning task was the same as in study 4, the
conditions of collaboration (viewing multiple maps or not) and the conditions
(scripted or not, awareness tool or not) probably explain differences in terms
of mutual modeling

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrv}
\bibliography{biblio}

\end{document}