### 2016-02-04 rebuttal.tex
WIP on the rebuttal

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript and providing in-depth review of our submission.

The main points that were raised relate to:
\begin{itemize}
    \item the need for a better presentation of the concepts that we build upon,
    \item the need for a ``deeper synthetic discussion'', also ``accounting for the
        conflicting results'', that would present the ``theoretical
        picture'' resulting from our work,
    \item discuss a possible research agenda
\end{itemize}

In line with these comments and suggestions, the manuscript underwent several
significant changes that we summarize hereafter.

\begin{itemize}

    \item The title and the abstract have been changed to clarify the
        claims of the paper, which, as the editor pointed out, were not clear
        for some reviewers. In a nutshell, we do not know what is going on in the
        brain when building a shared solution (we leave this scientific debate
        to psycholinguists), but we observe how they vary (or not) across
        situations (across the 5 experiments) and this enables some reflexions
        that we put in the context of collaborative learning.

    \item \emph{Section 1 -- Introduction} We did minor edits except for the last
        paragraph has been completely rewritten it  that was somehow misleading
        regarding to the goals of the paper.  We think the point is now made
        clearly that the issue is not whether it is alway necessary to build a
        shared understanding, but that CSCL folds design tasks specifically for
        requiring shared understanding.

    \item We clarified the notion of grounding criterion, as requested by
        reviewer 2 (page [TODO!!])

    \item \emph{Section 2} has been split into partner modeling and mutual modeling (now
        section 3) to emphasize the fact that the question is the mutuality. The
        main research question is then clearly stated to address some [WHICH?] of the
        reviewers' criticisms.

    \item \emph{Section 'Notations'} We clarified the status of the proposed formalism, which some
        reviewers found useful and others criticized. We are not
        proposing here a computational theory of mutual modeling. This notation is
        only useful because the text become quickly unreadable as one talks about
        the model that A build about the model that B build about A, while
        M(A,B,M(B,A,X)) is less ambiguous. We mainly see it as a language for discussing
        experiments and hypotheses in a simple way. Its function is internal to
        this paper.

    \item \emph{Section 'Studies'} For all the 5 studies, we have re-written all
        the \emph{discussion} sub-sections in a
        way that they naturally align with the research questions, leading to a
        clearer overall message.

    \item As requested by the reviewers, we have entirely rewritten  the
        conclusions to make them more explicit.  We cannot of course not pretend
        that our results are clearcut. We have to be careful not to publish only
        papers with significant results (statisticians have proved that some
        meta-analyses are biased by the fact that papers with significant results
        get more easily published). However, despite the fact that none of the 3
        hypothesis is declared as winner, the key results is that we prove that
        individual skills only explain a part of the variance, that the quality of
        social interactions matter significantly.

\end{itemize}

\end{document}